---
title: "Getting Started with UtilsCytoRSV"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with UtilsCytoRSV}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

UtilsCytoRSV is an R package that provides utility functions for working with cytometry data, including CyTOF (mass cytometry) and flow cytometry. This vignette provides an overview of the main functionality.

```{r setup}
library(UtilsCytoRSV)
```

## Visualization

### Basic Plotting with `plot_cyto()`
The `plot_cyto()` function creates 2D hex plots with useful defaults for cytometry data.

```{r plot-cyto-example, fig.height=4, fig.width=6, eval=requireNamespace("flowCore", quietly = TRUE)}
suppressWarnings(data("GvHD", package = "flowCore"))
ex_tbl <- flowCore::exprs(GvHD[[1]]) |>
  as.data.frame()
marker <- c("FL2-H", "FL3-H")
plot_cyto(
  data = ex_tbl,
  marker = marker
)
```

### Channel/Marker Labeling
Use `chnl_lab()` to get a named vector that maps channel names to marker names from an FCS file:

```{r chnl-lab-example, eval=requireNamespace("flowCore", quietly = TRUE)}
lab_vec <- chnl_lab(GvHD)
head(lab_vec)
```

## Calculations

### Frequencies and Proportions

Calculate frequencies (percentage) or proportions from numerator and denominator columns using `calc_freq()` and `calc_prop()`:

```{r calc-prop-example}
mock_data <- data.frame(
  pop = "cd4",
  cd4 = c(2000, 2100, 1900, 2050),
  ifng = c(500, 520, 480, 510)
)

# Calculate frequency (percentage)
result <- calc_freq(
  .data = mock_data,
  den = "cd4",
  num = "ifng"
)
result$freq
```

## Data Processing

### Background Subtraction

The `subtract_background()` function subtracts the unstimulated measurement from response columns:

```{r subtract-bg-example}
data_test <- data.frame(
  pid = rep(c("a", "b"), each = 3),
  stim = c("mtb", "ebv", "uns", "uns", "ebv", "mtb"),
  resp1 = 1:6,
  resp2 = c(34, 32, 30, 28, 26, 24)
)

data_out <- subtract_background(
  .data = data_test,
  grp = "pid",
  stim = "stim",
  uns = "uns",
  resp = c("resp1", "resp2"),
  remove_uns = FALSE
)
data_out
```

### Summing Over Markers

The `sum_over_markers()` function aggregates proportions or frequencies over specified markers:

```{r sum-markers-example}
data("data_count")
data_test <- data_count |>
  calc_prop(
    den = "count_pop_den",
    num = "count_pop_num"
  ) |>
  dplyr::select(-c(count_pop_den, count_pop_num)) |>
  dplyr::arrange(SubjectID, VisitType, stim, cyt_combn)

data_out <- sum_over_markers(
  .data = data_test,
  grp = c("SubjectID", "VisitType", "stim"),
  cmbn = "cyt_combn",
  markers_to_sum = c("IFNg", "IL2", "IL17"),
  levels = c("-", "+"),
  resp = "prop"
)
head(data_out)
```

## Session Info

```{r session-info}
sessionInfo()
```
